---
title: "TaxPub XML to Turtle RDF Converter"
author: "Viktor Senderov"
date: "`r Sys.Date()`"
---

# Specification

## Making the scripts

To create the script:
```
knitr::knit('exec/taxpub2rdf.Rmd', output = 'exec/taxpub2rdf.R', tangle = TRUE)
```

To create the documentation:
```
knitr::knit('exec/taxpub2rdf.Rmd', output = 'exec/taxpub2rdf.R', tangle = TRUE)
```

## Summary

`taxpub2rdf` is a command-line tool, written in the [R programming language](https://cran.r-project.org/), which RDF-izes, i.e. converts to [Resource Description Framework](https://www.w3.org/RDF/) (RDF) taxonomic papers, encoded in the [TaxPub XML schema](http://htmlpreview.github.io/?https://github.com/tcatapano/TaxPub/blob/master/documentation/tp-taxon-treatment.html). It is designed to be used together with the [Open Biodiversity Knowledge Management System](http://openbkms.blogspot.com) (OBKMS) graph database, as before it generates new identifiers for things in the network, it tries to look them up in OBKMS.

## Usage and requirements

`taxpub2rdf` requires two parameter-sets for successful operation:

1. Access details for OBKMS, which are stored as a Yaml file, except for the username-password combination, which is stored in the environment variable OBKMS_SECRET.
2. A set of input XML files in the TaxPub format.

General syntax:
```
Rscript taxpub2rdf.R config.yml paper1.xml paper2.xml ...
```

The first argument is the Yaml configuration file. The next n arguments are the TaxPub XML files, that will be RDF-ized. The output will be n Turtle files with the same filenames but the `.ttl` extension.

Example usage:
```
Rscript exec/taxpub2rdf.R etc/config.yml Standard_Model_Conversions/10.3897-BDJ.4.e10095.xml
```

# Design and Implementation

## Initialization

The script relies on the `xml2` library for XML processing and on the `rdf4jr` library for communication with the GraphDB server:

```{r , purl=TRUE, eval = FALSE}
library(xml2)
library(rdf4jr)]
source(R/query.R)
```

Read in the command-line arguments:

```{r , purl=TRUE, eval = FALSE}
command_line_args = commandArgs(trailingOnly = TRUE)
```

We have to have at least two command line arguments:

```{r , purl=TRUE, eval = FALSE}
stopifnot( length( command_line_args ) >= 2 )
```

The first argument is the Yaml configuration file, while the rest n - 1 arguments are the TaxPub's to be processed.

```{r , purl=TRUE, eval = FALSE}
configuration_file = command_line_args[ 1 ]
configuration_file = "/media/viktor/www2/main/vsenderov.backup/SRC/TaxPub2RDF/etc/config.yml"
taxpubs = command_line_args[ 2:length( command_line_args ) ]
options = yaml::yaml.load_file( configuration_file )
#taxpubs = c("/media/viktor/www2/main/vsenderov.backup/SRC/TaxPub2RDF/Standard_Model_Conversions/10.3897-BDJ.4.e10095.xml")
```

Get the username-password or API key from the environment:

```{r , purl=TRUE, eval = FALSE}
options$userpwd = Sys.getenv("OBKMS_SECRET")
options = c( options, obkms::create_options() )
```

Test GraphDB connectivity:
```{r, purl = TRUE, eval = FALSE}
stopifnot( get_protocol_version( obkms_options ) != 6 )
```

Read in the XML file:
```
### currently only one file accepted
taxpub_xml = read_xml(taxpubs)
```

Now, we will write different functions to extract different features of the XML.

```

```

We write a function that returns RDF about the publisher. RDF will be returned as strings in Turtle Syntax.
```
get_publisher_rdf = function(taxpub_xml) {
  publisher_name = xml_text(xml_find_all(taxpub_xml, "/article/front/journal-meta/publisher/publisher-name")[[1]])
  publisher_id = get_or_create_node_by_label( obkms_opts, publisher_name) 
  if (publisher_id == null) publisher_id = mint_new_pensoft_id()
  
  journal_name = xml_text(xml_find_all(taxpub_xml, "/article/front/journal-meta/journal-id")[[1]])
  journal_id = get_or_create_node_by_label( obkms_opts, journal_name)
  
  
  rdf = "
  prefix fabio: <fdfsdsfsdfs> .
  %journal_id a fabio:Journal"
  return(rdf)
}
```

# this is in semi pseudo code
extract_journal_metadata = function(y_document) {
  journal_name = get_xml_value("/article/front/journal-meta/journal-id")
}
