#### Example: *Article Metadata*

The main objects of information extraction and retrieval of OpenBiodiv in the
first stage of its development are scientific journal articles from the
journals [Biodiversity Data Journal](http://bdj.pensoft.net/) and
[ZooKeys](http://zookeys.pensoft.net/) and other Pensoft journals.
We model the bibliographic objects around Journal Article, such as Publisher,
and Journal using SPAR.

Note that we type `:treatment` both as `trt:Treatment` (i.e. the rhetorical
element Treatment) and as s `doco:Section` because we view this particular
treatment to also be a structural section of the document.




```{r Example Article Metadata}
:biodiversity-data-journal rdf:type fabio:Journal ;
	 skos:prefLabel "Biodiversity Data Journal"@en ;
	 skos:altLabel  "BDJ"@en ;
	 fabio:issn     "1314-2836" ;
	 fabio:eIssn    "1314-2828" ;
	 frbr:part      :b90f6933-ab5e-4ce1-9379-12de9ef4eaa6 . 

 <http://dx.doi.org/10.3897/BDJ.1.e953> rdf:type fabio:TaxonomicArticle ;
	 skos:prefLabel           "10.3897/BDJ.1.e953" ;
	 dc:title                 "Casuarinicola australis Taylor, 2010 
         (Hemiptera: Triozidae), newly recorded from New Zealand"@en ;
	 prism:doi                "10.3897/BDJ.1.e953" ;
	 dcelements:publisher     "Pensoft Publishers"@en ;
	 fabio:hasPublicationYear "2013"^^xsd:gYear ;
	 prism:publicationDate    "2013-9-16"^^xsd:date ;
	 dcterms:publisher        :pensoft-publishers ;
	 frbr:realizationOf       :thorpe-2013 .

 :thorpe-2013 rdf:type  :ResearchPaper ;
 	 skos:prefLabel	      "Thorpe 2013"
	 skos:altLabel        "paper10.3897/BDJ.1.e953" ;
	 dcterms:creator      :stephen-e-thorpe ;
	 prism:keywords       "Casuarinicola australis"@en ;
	 fabio:hasSubjectTerm :a2ee4929-90dd-4a7a-aa5c-08836f49d549 . 

 :pensoft-publishers rdf:type :Publisher ;
	 skos:prefLabel "Pensoft Publishers"@en . 

 :stephen-e-thorpe rdf:type foaf:Person ;
	 skos:prefLabel "Stephen E. Thorpe" ;
	 foaf:firstName "Stephen E." ;
	 foaf:surname   "Thorpe" ;
	 foaf:mbox      "stephen_thorpe@yahoo.co.nz" ;
	 :affiliation   "School of Biological Sciences (Tamaki Campus),
	       University of Auckland, Auckland, New Zealand"@en . 

 :a2ee4929-90dd-4a7a-aa5c-08836f49d549 rdf:type fabio:SubjectTerm ;
	 rdfs:label    "Casuarinicola australis"@en ;
	 skos:inScheme :openbiodiv-subject-terms . 
```


#### Example: *Article Structure*

```{r Example Article Structure}
<http://dx.doi.org/10.3897/BDJ.1.e953>
  :contains :abstract, :casuarinicola-australis-treatment .

:introduction rdf:type deo:Introduction, doco:Section ;
  c4o:hasContent "Casuarinicola australis Taylor, 2010 was described from
        Australia, where it is the most common and widespread member of its
        genus, being widely distributed in New South Wales, Queensland,
        South Australia, Victoria and Western Australia. "

:casuarinicola-australis-treatment rdf:type doco:Section, :Treatment ;
  :contains :casuarinicola-australis-nomenclature ,
              :casuarinicola-australis-materials ,
              :casuarinicola-australis-description ,
              :figure-box-1 ,
              :figure-box-2 .

:casuarinicola-australis-nomenclature rdf:type :NomenclatureSection ;
  :contains :casuarinicola-australis-nomenclature-heading .

:casuarinicola-australis-nomenclature-heading a :NomenclatureHeading ;
  cnt:chars "Casuarinicola australis Taylor, 2010" .

:casuarinicola-australis-materials rdf:type :MaterialsExamined ;
  c4o:hasContent "country: New Zealand;
                  verbatimLocality: Mechanics Bay, Auckland City;
                  verbatimElevation: 0-5 m;
                  verbatimLatitude: 36.8474938105S ;
                  verbatimLongitude: 174.7869624545E ;
                  eventDate: 6 January 2013;
                  sex: 1 male, 1 female;
                  recordedBy: Stephen Thorpe;
                  institutionCode: Auckland Museum" .

:casuarinicola-australis-description rdf:type :DescriptionSection ;
  c4o:hasContent "On 6 Jan 2013, I examined some Casuarina glauca trees growing
        in the vicinity of Ports of Auckland at Mechanics Bay." .
```




#### Example

In the following example, we express in RDF a TNU that is in the
nomenclature heading of a treatment (treatment title). Structurally, the TNU
is connected to the containing section via `po:contains`; `cnt:chars` is used
to dump the full string of the usage and DwC properties are used to encode
more granular information in addition to the dump.

In the second step of RDF-ization, we use `dwciri` properties to link the TNU
to semantic entities. `dwciri:taxonomicStatus` is used to link the TNU to an
item in the
[OpenBiodiv Taxonomic Status Vocabulary](#vocabulary-of-taxon-classification).
`:scientificName` is used to link the TNU to the IRI of the name that
the TNU is mentioning. Note, we have introduced `:scientificName` as
a sub-property of `pext:Mention`. In this example it is linked both to local
name and to a remote name. This implies that the names are the same (see Rule
later).

Also, during the second step, the TNU is linked to the reified taxon concept
label *Heser stoevi* sec. 10.3897/BDJ.4.e10095 via `:taxonConceptLabel` as
even though the character content of the TNU does not contain an "according
to" (usually abbreviated as "sec."), we know for certain which concept the
author is invoking as we are in the treatment title (current concept/ *this*
concept).

```# Example: Taxonomic Name Usage

:casuarinicola-australis-nomenclature-heading
  po:contains :casuarinicola-australis-TNU .

:casuarinicola-australis-TNU a :TaxonomicNameUsage ;
  dc:date "2013-9-16"^^:xsd:date ;
  cnt:chars "Casuarinicola australis Taylor, 2010" ;
  dwc:genus "Casuarinicola" ;
  dwc:specificEpithet "australis" ;
  dwc:scientificNameAuthorship "Taylor, 2010" ;
  # we can infer the following because we are in the treatment heading
  dwc:nameAccordingToId "doi: 10.3897/BDJ.1.e953" ;
  pkm:mentions :casuarinicola-australis-taylor, 
  			   :casuarinicola-australis-taylor-sec-thorpe-2013 .

:casuarinicola-australis-taylor a :ScientificName ;
  rdfs:label "Casuarinicola australis Taylor, 2010" ;
  dwc:genus "Casuarinicola" ;
  dwc:specificEpithet "australis" ;
  dwc:scientificNameAuthorship "Taylor, 2010" .

:casuarinicola-australis-taylor-sec-thorpe-2013 a :TaxonomicConceptLabel ;
  rdfs:label "Casuarinicola australis Taylor, 2010 sec. Thorpe 2013" ;
  dwc:genus "Casuarinicola" ;
  dwc:specificEpithet "australis" ;
  dwc:scientificNameAuthorship "Taylor, 2010" .
  dwc:nameAccordingToId "doi: 10.3897/BDJ.1.e953" ;
  :nameAccordingTo <http://dx.doi.org/10.3897/BDJ.1.e953> .
```

```{r Example Simple Taxonomic Concept Relationships}
:concept-casuarinicola-australis-thorpe rdf:type :TaxonomicConcept ;
  :taxonomicConceptLabel :casuarinicola-australis-taylor-sec-thorpe-2013 .

:concept-casuarinicola-taylor rdf:type :TaxonomicConcept ;
  skos:broader concept-thorpe .
```

```{r Example Simple Taxonomic Concept Relationships}
:statement rdf:type :RCC5Statement ;
  :rcc5FromRegion :concept-casuarinicola-australis-thorpe ;
  :rcc5ToRegion :concept-casuarinicola-taylor ;
  :rcc5AccordingTo <http://dx.doi.org/10.3897/BDJ.1.e953> ;
  :rcc5RelationType :ProperPart_INT .
```

```{r Example ENVO}
:australian-casuarina-forest rdf:type <http://purl.obolibrary.org/obo/ENVO_01000174> .
:hasHabitat owl:sameAs <http://purl.obolibrary.org/obo/RO_0002303> .
:concept-casuarinicola-australis-thorpe :hasHabitat :australian-casuarina-forest .
```

```{e Example Treatment - Concept}
:casuarinicola-australis-treatment frbr:realizationOf :concept-casuarinicola-australis-thorpe.
```

```
<<Examples>>=

:heser-stoevi-nomenclature-heading po:contains :heser-stoevi-TNU .

:heser-stoevi-TNU a :TaxonomicNameUsage ;
  dc:date "2016-08-31"^^xsd:date ;
  cnt:chars
  "Heser stoevi urn:lsid:zoobank.org:act:E4D7D5A0-D649-4F5E-9360-D0488D73EEE8 Deltschev sp. n." ;
  dwc:genus "Heser" ;
  dwc:species "stoevi" ;
  dwc:scientificNameId "urn:lsid:zoobank.org:act:E4D7D5A0-D649-4F5E-9360-D0488D73EEE8 (ZooBank)" ;
  dwc:scientificNameAuthorship "Deltschev" ;
  dwc:taxonomicStatus "sp. n." ; dwciri:taxonomicStatus :TaxonDiscovery ;
  dwc:nameAccordingToId "10.3897/BDJ.4.e10095" ;

  :scientificName :heser-stoevi-deltshev ;
  :nameAccordingTo <http://dx.doi.org/10.3897/BDJ.4.e10095> ;
  :taxonConceptLabel :heser-stoevi-sec-deltshev ;

  :scientificName <http://zoobank.org/urn:lsid:zoobank.org:act:E4D7D5A0-D649-4F5E-9360-D0488D73EEE8> .

:heser-stoevi-deltshev owl:sameAs <http://zoobank.org/urn:lsid:zoobank.org:act:E4D7D5A0-D649-4F5E-9360-D0488D73EEE8> .
@
     
```



**Example** We go back to the example of  *Heser stoevi*. The meaning of the
date property here is to indicate when was the taxonomic status assumed.

```
<<Examples>>= 

:heser-stoevi-deltshev a :ScientificName ;
	skos:prefLabel "Heser stoevi Deltshev" ;
    dwc:species "stoevi" ;
    dwc:genus "Heser" ;
    dwc:taxonRank "species" ;
    dwciri:taxonomicStatus <http://rs.gbif.org/vocabulary/gbif/taxonomicStatus/accepted> ;
    dwc:scientificNameAuthorship "Deltschev" ;
    dc:date "2016-08-31"^^xsd:date .
@
```

**Example.** Let's take another example, the paper
<http://bdj.pensoft.net/articles.php?id=8030&instance_id=2809105>. From it, we
can say:

```
<<Examples>>= 

:nomenclature-bdje8030 a trt:Nomenclature ;
  po:contains :nomenclature-heading-bdje8030, :cit-list-bdje8030 .

:nomenclature-heading-bdje8030 a :NomenclatureHeading ;
  po:contains :harmonia-manillana-tnu-heading .

:cit-list-bdje8030 a trt:NomenclatureCitationList ;
  po:contains :leis-papuensis-tnu-citation .

:harmonia-manillana-tnu-heading a :TaxonomicNameUsage ;
  dc:date "2016-08-16"^^xsd:date ;
  cnt:chars "Harmonia manillana (Mulsant, 1866)" .

:leis-papuensis-tnu-citation a :TaxonomicNameUsage ;
  dc:date "2016-08-16"^^xsd:date ;
  cnt:chars "Leis papuensis var. suffusa Crotch 1874 121 (Lectotype, UCCC). Korschefsky 1932 : 275.- Gordon 1987 : 14 (lectotype designation). Syn. nov." ;
  dwc:taxonomicStatus "var. suffusa Crotch 1874 121 (Lectotype, UCCC). Korschefsky 1932 : 275.- Gordon 1987 : 14 (lectotype designation). Syn. nov." ;
  dwciri:taxonomicStatus :Unavailable .

:harmonia-manillana-mulsant-1866 a :ScientificName ;
  skos:prefLabel "Harmonia manillana (Mulsant, 1866)" ;
  skos:altLabel "Harmonia manillana" ;
  dwc:species "manillana" ;
  dwc:genus "Harmonia" ;
  dwc:taxonRank "species" ;
  dwc:scientificNameAuthorship "(Mulsant, 1866)" ;
  dc:date "2016-08-16"^^xsd:date ; 
  dwc:taxonomicStatus :Available ;
  :relatedName :leis-papuensis .

:leis-papuensis a :ScientificName ;
  skos:prefLabel "Leis papuensis" ;
  dwc:species "papuensis" ;
  dwc:genus "Leis" ;
  dwc:taxonRank "species" ;
  dc:date "2016-08-16"^^xsd:date ; 
  dwciri:taxonomicStatus :Unavailable ;
  :replacementName :harmonia-manillana-mulsant-1866 ;
  :relatedName :harmonia-manillana-mulsant-1866 .
@
```


**Example.** In the next example we introduce the concept of *Heser stoevi*
according to the article published by Deltshev in 2016. First, we introduce an
instance of `:TaxonConcept` and link this instance to the scientific name
*Heser stoevi* via the appropriate DwC term. Next, we establish a link between
the significant bibliographic unit (in this case journal article) containing
the treatment, which is the realization of the taxon concept. The last point
we would like to make is that the taxon concept label, which is in this case
`Heser stoevi sec. 10.3897/BDJ.4.e100095` is constructed by pasting together
the label of the biological name and the expression that are assigned to the
concept glued together by `sec.`.

```
<<Examples>>=

:heser-stoevi-sec-deltshev a :TaxonConceptLabel ;
   skos:prefLabel "Heser stoevi Deltshev sec. 10.3897/BDJ.4.e10095" ;
   dwc:species "stoevi" ;
   dwc:genus "Heser" ;
   dwc:taxonRank "species" ;
   dwc:scientificNameAuthorship "Deltschev" ;
   ::nameAccordingTo <http://dx.doi.org/10.3897/BDJ.4.e10095> .
  
:concept-deltshev-2016 a :TaxonConcept ;
  :taxonConceptLabel :heser-stoevi-sec-deltshev .

:heser-stoevi-sec-gbif20170323 a :TaxonConceptLabel ;
  skos:prefLabel "Heser stoevi sec. doi:10.15468/39omei" ;
  dwciri:scientificName :heser-stoevi-deltshev ;
  dwciri:nameAccordingTo :gbif20170323 .

:concept-gbif a :TaxonConcept ;
  :taxonConceptLabel :heser-stoevi-sec-gbif20170323 .

<http://dx.doi.org/doi:10.15468/39omei> a fabio:Database ;
  skos:prefLabel "GBIF Backbone Taxonomy" ;
  skos:altLabel "doi:10.15468/39omei" ;
  prism:doi "doi:10.15468/39omei" ; 
  dc:date "2017-03-23"^^xsd:date ;
  rdfs:comment "A dump of GBIF's backbone taxonomy on 23 Mar 2017."@en ;
  po:contains [ a :TaxonConceptLabel ;
                dc:date "2017-03-23"^^xsd:date ;
                dwc:scientificName "Heser stoevi Deltshev, 2016" ;
                dwc:nameAccordingTo "GBIF Backbone Taxonomy" ;
                dwciri:nameAccordingTo <http://dx.doi.org/doi:10.15468/39omei> ;
                dwciri:scientificName :heser-stoevi-deltshev ;
                pkm:mentions :heser-stoevi-sec-gbif20170323  ] .
@
```

Note that in the above example one scientific name, *Heser stoevi*, is linked
to two different taxon concept labels, as one taxon concept label denotes the
concept coming from the article and the other one comes from the GBIF
database.

It is possible to express that these are the same thing, that one is a
subconcept of the other, or even more granular relationships.


#### Taxon Concept Relationships

**Example congruence 1.** If we want to express that two concepts
are exactly the same both intensionally and ostensively, then we use `owl:sameAs`.
However, if we want to express that the taxon concepts have the same extension
without being equal intensionally (as in the spider example), then we use
`owl:equivalentClass`.

```
<<Examples>>=

:heser-stoevi-sec-deltschev owl:sameAs :heser-stoevi-sec-gbif20170323 .
@
```

Note that this will copy both taxon concept labels both ways, but this is OK
as the different taxon concept labels refer to the same class.

**Example congruence 2.** Let's define the two spider concepts here.

```
<<Examples>>=

:haveSpinnerets rdf:type owl:DatatypeProperty ;
  rdfs:domain :TaxonConcept ; 
  rdfs:comment "This property if true, indicates that some taxon has spinnertes." .

:havePedipals rdf:type owl:DataTypeProperty ;
  rdfs:domain :TaxonConcept ;
  rdfs:comment "If this property is true, that the taxon has pedipals" .

:spiders-with-spinnerets-sec-rdfguide a :TaxonConceptLabel ;
  dwc:order "Araneae" ;
  skos:prefLabel "Araneae sec. OpenBiodiv RDF Guide" ;
  :nameAccordingTo <https://github.com/pensoft/OpenBiodiv/blob/master/Ontology/RDF_Guide.md> .

:spiders-with-pedipals-sec-rdfguide a :TaxonConceptLabel ;
  dwc:order "Aranea" ;
  skos:prefLabel "Araneae sec. OpenBiodiv RDF Guide" ;
  :nameAccordingTo <https://github.com/pensoft/OpenBiodiv/blob/master/Ontology/RDF_Guide.md> .

:concept-spinnerets a :TaxonConcept ;
  :taxonConceptLabel :spiders-with-spinnerets-sec-rdfguide ;
  :haveSpinnerrets "true"^^xsd:boolean .

:concept-pedipals a :TaxonConcept ;
  :taxonConceptLabel :spiders-with-pedipals-sec-rdfguide ;
  :havePedipals "true"^^xsd:boolean .

:spider1 a :concept-spinnerts .
:spider2 a :concept-pedipals .

:concept-spinnerts owl:equivalentClass :concept-pedipals .
@
```

Here, the implication is that although the intensional meaning of the two concepts
is different, they have the same class extension. I.e. we will infer

```
:spider1 a :concept-pedipals;
:spider2 a :concept-spinnerts'
```

but we will not copy any of the `:havePedipals` or `:haveSpinnerts` to the other concept.

**Example of contained concepts.** For contained concept we use `rdfs:subClassOf`:

```
<<Examples>>=

:animalia-sec-gbif a :TaxonConceptLabel ;
  skos:prefLabel "Animalia sec. GBIF Backbone Taxonomy" ;
  dwciri:scientificName :animalia ;
  :nameAccordingTo <http://dx.doi.org/doi:10.15468/39omei> .

:animalia a :ScientificName ;
  dwc:scientificName "Animalia" ;
  dwc:taxonRank "kingdom" ;
  dc:date "2017-03-23"^^xsd:date ;
  dwciri:taxonomicStatus :Available .

:concept-animalia-gbif a :TaxonConcept ;
  dwciri:scientificName :animalia ;
  dwc:taxonId "1 (GBIF)" .

:heser-stoevi-sec-gbif20170323 rdfs:subClassOf :animlia-sec-gbif20170323 .
@
```

**Example (Relatedness)** If two taxon concepts are related we can use `skos:related`.

```
<<Examples>>=

:heser-nicola a :ScientificName ;
  dwc:scientificName "Heser nilicola (O. P.-Cambridge, 1874)" ;
  skos:prefLabel "Heser nilicola" ;
  dwc:genus "Heser" ;
  dwc:species "nilicola" ;
  dwc:taxonRank "species" .

:heser-nicola-sec-unibe a :TaxonConceptLabel ;
  skos:prefLabel "Heser nilicola (O. P.-Cambridge, 1874) sec. Unibe" ;
  dwciri:scientificName :heser-nicola ;
  frbr:realization <http://www.araneae.unibe.ch/data/3301> .

:concept-heser-nicola-unibe a :TaxonConcept ;
  :taxonConceptLabel :heser-nicola-sec-unibe .

:heser-stoevi-deltschev-sec-deltschev skos:related :heser-nicola-sec-unibe .
@
```


We model the following exceprt here:

"Minyomerus microps (Say, 1831: 9) sec. Jansen & Franz (2015), stat. n.  ==
(INT) AND > (OST) AND = Minyomerus innocuus Horn, 1876: 18 sec. Horn (1876) (type, designated by Pierce 1913: 400), syn. n."

```
<<Examples>>=

:jansen-franz-2015 a fabio:JournalArticle ;
  skos:prefLabel "10.3897/zookeys.528.6001" . 

:concept-minyomerus-microps-jansen-franz a :TaxonConcept .

:concept-minyomerus-innocuus-horn a :TaxonConcept .

:microps-innocuus-relation-int a :RCC5Stament ;
  :rcc5FromRegion :concept-minyomerus-microps-jansen-franz ;
  :rcc5ToRegion :concept-minyomerus-innocuus-horn ;
  :rcc5RelationType :Equals_INT ;
  frbr:expression :jansen-franz-2015 .

:microps-innocuus-relation-ost a :RCC5Stament ;
  :rcc5FromRegion :concept-minyomerus-microps-jansen-franz ;
  :rcc5ToRegion :concept-minyomerus-innocuus-horn ;
  :rcc5RelationType :InverseProperPart_OST ;
  frbr:expression :jansen-franz-2015 .
@
```
