# OpenBiodiv Ontology and Guide

OpenBiodiv expands to "Open Biodiversity Knowledge Management System"
and has previously been referred to as OBKMS.

## Preliminaries

This document is the guide to the OpenBiodiv ontology (OpenBiodiv-O) and
the ontology at the same time [as a literate programming]
(https://en.wikipedia.org/wiki/Literate_programming)
document. This means that we include the source code within its
documentation.

Previous versions relied on the program `noweb`. We have now switched to the R
library  `knitr` for building the document from Rmarkdown. Since there is no
Turtle engine for `knitr`, we markup the code-chunks as R, but do not execute
them.

The R `knitr` command to create the Turtle file with the ontology from this
document is:

```
knit(input = "openbiodiv-ontology.Rmd", tangle = TRUE, output = "openbiodiv-ontology.ttl")
```

It obtains the ontology in its Turtle serialization.

## Introduction

The present work introduces OpenBiodiv Ontology, which serves as the basis of
the OpenBiodiv knowledge system. By providing an ontology focusing on
taxonomy, our intent is to provide an ontology that fills in the gaps between
ontologies for biodiversity resources such as Darwin-SW and semantic
publishing ontologies such as SPAR. Moreover, we take the view that is
advantageous to model the taxonomic process itself, and not any particular
state of knowledge.

### Previous work

A lot of research has gone into ontologies, knowledge bases in general and
into biodiversity knowledge representation in particular. This gives us a vast
amount of publications, ontologies, vocabularies, and datasets to draw from
while implementing our model and database. In this section we list these
sources of inspiration.

#### What is an ontology

1. Obitko (2007) defines an ontology as a
[Specification of Conceptualization](https://www.obitko.com/tutorials/ontologies-semantic-web/specification-of-conceptualization.html)

2. Guarino et al. (2009) define an ontology a
[Ga shared formal, explicit specification of a conceptualization](http://iaoa.org/isc2012/docs/Guarino2009_What_is_an_Ontology.pdf).
This article goes into set-theoretic details of what is conceptualization and
formalisms are in order to properly write down a conceptualization in a
mathematical form.

#### Data models

3. A very important data model that we are drawing from is the
[Semantic Publishing and Referencing Ontologies][http://http://www.sparontologies.net/ontologies/]
by Peroni (2014). We use it to model scientific articles, their structure, and related
entities. As part of SPAR we use individual ontologies such as
[FaBiO](http://www.sparontologies.net/ontologies/fabio),
[DOCO](http://www.sparontologies.net/ontologies/fabio) and so on.

4. Another very important data model that we drawing from is the
[Darwin-SW](http://www.semantic-web-journal.net/system/files/swj635.pdf)
by Baskauf and Web (2014).

5. Furthermore we use the [Darwin Core RDF Guide](http://rs.tdwg.org/dwc/terms/guides/rdf/).

6. As a top-level ontologies we use [PROTON](http://ontotext.com/proton/) and
[SKOS](https://www.w3.org/2004/02/skos/).

7. For modeling scientific names we draw inspiration from [NOMEN](https://github.com/SpeciesFileGroup/nomen) 
and the 
[Taxonomic Nomenclatural Status Terms](https://github.com/plazi/TreatmentOntologies/blob/master/ontologies/taxonomic_nomenclatural_status_terms.owl_).

8. A specific part of a taxonomic manuscript called
["treatment" is modeled by Plazi](https://github.com/plazi/TreatmentOntologies).

9. An attempt to model taxon concepts has previously been made in the
[Taxon Concept Ontology](http://lod.taxonconcept.org/ontology/doc/index.html).

#### Concepts

10. [Open Biodiversity Knowledge Management System](http://riojournal.com/browse_user_collection_documents.php?collection_id=1&journal_id=17) is a PhD
project by Viktor Senderov. Rod Page
[has also published an article](http://riojournal.com/articles.php?id=8767)
on the same topic.

11. It is an attempt to model
[Taxonomic and Biodiversity Information for Computers](https://link.springer.com/article/10.1007/s13752-017-0259-5).

12. [Names ultimately have a limited use in informatics](http://zookeys.pensoft.net/articles.php?id=6234),
also [challenges](http://bdj.pensoft.net/articles.php?id=8080).
Original research in this work models them names and their relationships

13. It models
[potential taxa](https://link.springer.com/article/10.1007/s13752-017-0259-5)
as taxon concepts. Taxon concepts are also treated in 
[Two Influential Primate Classifications Aligned](https://academic.oup.com/sysbio/article/65/4/561/1753624/Two-Influential-Primate-Classifications-Logically) and 
[the phylogentic revision of the genus *Minyomerus*](http://zookeys.pensoft.net/articles.php?id=6001)
by Franz et. al (2015, 2016).

14. Taxon concepts can have both an intensional meaning and a class
extension. Some examples come from
[From Cladograms to Classifications](http://www.systass.org/archive/events-archive/2001/platnick.pdf)
by Platnick (2001).

14. The Codes of [Zoological](http://www.iczn.org/iczn/index.jsp) and
[Botanical](http://www.iapt-taxon.org/nomen/main.php) Nomenclature ought also
to be mentioned as a source of albeit too granular in some cases inspiration
for the data models.

#### Semiotics

14. When we model the real world we always run up against the sign theories
of [Frege](https://en.wikipedia.org/wiki/Triangle_of_reference) and of
[Pierce](https://plato.stanford.edu/entries/peirce-semiotics/).

### Note on Identifier Naming

Our ontology strives be a formal specification of a conceptualization. In our
mental model we have some concepts of some things. When we talk about these
concepts in the abstract, we will make use of Capitalization. For example
let's suppose we want to name the top-level concept, i.e. the concept which
encompasses all  concepts. For this we write Thing. In another example, let's
suppose we want to name a part of a manuscript where taxonomic circumscription
takes place. We say Treatment when we refer to that concept. We also have
concepts for relations (in our conceptualization only binary relations are
allowed). To denote these relations in the abstract we use verbal phrases and
we might or might not use quotes (we will use quotes only if it adds to the
clarity of exposition). For example, Treatment is a Thing as opposed to
Treatment "is a" Thing. We also have individual instances of these concepts.
To refer to those we might use improper or proper nouns or phrases wherever
appropriate. For example, "the treatment on page 5," or "a treatment," or
"John."

When we formally define a concept in OWL and issue an URI to it, we shall
refer to the URI, as we refer to all URI's in the text with `typewriter font`.
URI's of classes and vocabularies will be in `MajorCamelCase`. URI's of
relationships will be in `minorCamelCase`. URI's of individuals `will-be-
hyphenated`. This seems to generally in accordance with WWW practice.

Also we strive for human-readable identifiers even if some ontology building
practices argue against it.

## Ontology

### Setup

##### Prefixes

*Note*: As SPAR (and FaBiO in particular) serve as somewhat of a _top-level_
ontology for us, we use the `skos` prefix as they use it.

```{r Prefix}
@prefix : <http://openbiodiv.net/> .
@prefix skos: <http://www.w3.org/2004/02/skos/core#> .
@prefix rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#> .
@prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#> .
@prefix foaf: <http://xmlns.com/foaf/0.1/> .
@prefix pro: <http://purl.org/spar/pro/> .
@prefix scoro: <http://purl.org/spar/scoro/> .
@prefix ti: <http://www.ontologydesignpatterns.org/cp/owl/timeinterval.owl#> .
@prefix tvc: <http://www.essepuntato.it/2012/04/tvc/> .
@prefix xsd: <http://www.w3.org/2001/XMLSchema#> .
@prefix fabio: <http://purl.org/spar/fabio/> .
@prefix dcelements: <http://purl.org/dc/elements/1.1/> .
@prefix dcterms: <http://purl.org/dc/terms/> .
@prefix frbr: <http://purl.org/vocab/frbr/core#> .
@prefix prism: <http://prismstandard.org/namespaces/basic/2.0/> .
@prefix doco: <http://purl.org/spar/doco/> .
@prefix po: <http://www.essepuntato.it/2008/12/pattern#> .
@prefix co: <http://purl.org/co/> .
@prefix trt: <http://plazi.org/treatment#> .
@prefix c4o: <http://purl.org/spar/c4o/> .
@prefix dwciri: <http://rs.tdwg.org/dwc/iri/> .
@prefix nomen: <http://www.semanticweb.org/dmitriev/ontologies/2013/8/untitled-ontology-6#> .
@prefix dwc: <http://rs.tdwg.org/dwc/terms/> .
@prefix sro: <http://salt.semanticauthoring.org/ontologies/sro#> .
@prefix deo: <http://purl.org/spar/deo/> .
@prefix pext: <http://proton.semanticweb.org/protonue#> .
@prefix ptop: <http://proton.semanticweb.org/protont#> .
@prefix pkm: <http://proton.semanticweb.org/protonkm#> .
@prefix vcard: <http://www.w3.org/2006/vcard/ns#> .
@prefix dbr: <http://dbpedia.org/resource/> .
@prefix org: <http://www.w3.org/ns/org#> .
@prefix owl: <http://www.w3.org/2002/07/owl#> .
@prefix lucene: <http://www.ontotext.com/connectors/lucene#> .
@prefix inst: <http://www.ontotext.com/connectors/lucene/instance#> .
@prefix dbo: <http://dbpedia.org/ontology/> .
@prefix vs: <http://www.w3.org/2003/06/sw-vocab-status/ns#> .
@prefix dwcattributesns: <http://rs.tdwg.org/dwc/terms/attributes/> .
```

##### Metadata

```{r Ontology Metadata}
:openbiodiv-ontology rdf:type owl:Ontology ;
  owl:versionInfo "1.0.1" ;
  rdfs:comment "You are free to create derivative work as long as you cite the original authors." ;
  dcelements:title "The Open Biodiversity Knowledge Management System Ontology" ;
  dcelements:subject "Biodiversity Informatics" ;
  rdfs:label "OpenBiodiv Ontology" ;
  dcelements:creator "Viktor Senderov, Kiril Simov" ;
  dcelements:license "CC-BY license version 4.0" ;
  owl:imports <http://purl.org/spar/fabio/> .
```

#### Imports and Contributions

1. We incorporate several external ontologies fully into our data model. Some of
them via a direct web import. Where it is not possible due to an error, we 
import them as text files found under [`imported_ontologies`](./imported_ontologies).
They may contain small bug-fixes to overcome the error.

2. We also borrow parts from external ontologies (contributed parts). Those are
attributed in the text.

### Semantic Model of the Publishing Domain

#### Discussion 

The publishing domain is described in our model using the Semantic Publishing
and Referencing Ontologies, a.k.a. [SPAR Ontologies](http://www.sparontologies.net/).
We do import several of these ontologies (please consult the paragraph
["Imports and Contributions"](#imports-and-contributions)).
Refer to the documentation on the SPAR Ontologies' site for an exhaustive treatment.

In the rest of this section we describe the modeling of entities in the
publishing domain that are *not found* in the SPAR ontologies. The central new
class in OpenBiodiv not found in SPAR is the `:Treatment` class, borrowed
from the [Treatment Ontologies](https://github.com/plazi/TreatmentOntologies).

##### Axiom

`frbr:realizationOf` is the inverse of `frbr:realization`

```{r}
frbr:realization owl:inverseOf frbr:realizationOf.
```

##### Contributed Properties: *contains*

```{r http://www.essepuntato.it/2008/12/pattern#contains}
po:contains rdf:type owl:ObjectProperty ;
          owl:inverseOf po:isContainedBy ;
          rdfs:domain po:Structured ;
          rdfs:comment "A structured element contains another generic element."@en ;
          rdfs:label "contains"@en .
```

```{r  http://www.essepuntato.it/2008/12/pattern#isContainedBy}
po:isContainedBy rdf:type owl:ObjectProperty ;
               rdfs:comment "An element that is contained by another structured element."@en ;
               rdfs:label "is contained by"@en .
```

##### Contributed Classes: *structured element*, *pattern-based element*

```{r http://www.essepuntato.it/2008/12/pattern#Structured}
po:Structured rdf:type owl:Class ;
            rdfs:subClassOf po:Pattern ;
            rdfs:comment "The class of elements that can contain other elements."@en ;
            rdfs:label "structured element"@en .
```

```{r http://www.essepuntato.it/2008/12/pattern#Pattern}
po:Pattern rdf:type owl:Class ;
         rdfs:subClassOf owl:Thing ;
         rdfs:comment "A structural pattern is an atomic component of a document, that is able to capture a recurrent and well known structure of a document. Any entity belonging to this class is compliant with exactly one structural pattern."@en ;
         rdfs:label "pattern-based element"@en .
```

##### Class: *Mention*

```{r http://proton.semanticweb.org/protonue#Mention}
:Mention rdf:type owl:Class ;
  rdfs:seeAlso pext:Mention ;
  rdfs:subClassOf deo:DiscourseElement ;
  rdfs:comment "An area of a document that can be considered a mention of something."@en ;
  rdfs:label "Mention" .
```    

##### Contributed Class: *Statement*

```{r}
ptop:Statement rdf:type owl:Class ;
               rdfs:comment "A message that is stated or declared; a communication (oral or written), setting forth particulars or facts, etc; \"according to his statement he was in London on that day\". WordNet 1.7.1"@en ;
               rdfs:label "Statement"@en .
``` 

##### Contributed Class: *Agent*

```{r http://xmlns.com/foaf/0.1/Agent}
foaf:Agent rdf:type owl:Class ;
           rdfs:comment "An agent (eg. person, group, software or physical artifact)." ;
           rdfs:label "Agent" ;
           vs:term_status "stable" .
```

##### Property: *mentions*

```{r http://proton.semanticweb.org/protonkm#mentions}
:mentions
      rdf:type owl:ObjectProperty ;
      rdfs:comment "A direct link between an information resource, like a document or a section and an entity." ;
      rdfs:domain :Mention ;
      rdfs:label "mentions" .
```


##### Property: *contains*

`contains` is a sub-property of `po:contains` that is transitive property.

```{r http://openbiodiv.net/contains}
:contains rdf:type owl:ObjectProperty, owl:TransitiveProperty ;
  rdfs:label "OpenBiodiv contains" ;
  rdfs:subPropertyOf po:contains. 

:isContainedBy rdf:type owl:ObjectProperty, owl:TransitiveProperty ;
  rdfs:label "OpenBiodiv is contained by" ;
  rdfs:subPropertyOf po:isContainedBy. 

```

We have mentioned before that when we extract bibliographic elements from the XML,
we make use of the `po:contains` SPAR property. For example, an article can 
`po:contain` a secion and this section can `po:contain` another (sub-)section.
In our view, this means that also the article contains the (sub-)section. Therefore
we define `po:contains` as a transitive property.

##### Class: *Publisher*

The publisher of a journal, a type of `foaf:Agent`.

```{r http://openbiodiv.net/Publisher}
:Publisher rdf:type owl:Class ;
  rdfs:label "Publisher"@en ;
  rdfs:comment "The publisher of a journal, a type of `foaf:Agent`."@en ;
  rdfs:subClassOf foaf:Agent .
```

##### Class: *Taxonomic Article*

`:TaxonomicArticle` is a specialized journal article for publishing taxonomic findings.

```{r http://openbiodiv.net/TaxonomicArticle}
:TaxonomicArticle rdf:type owl:Class ;
  rdfs:subClassOf fabio:JournalArticle ;
  rdfs:label "Taxonomic Article"@en .
```

##### Class: *Taxonomic Treatment*

See [Plazi](http://plazi.org/) for a theoretical discussion of Treatment.

A treatment is a rhetorical element of a taxonomic article. Thus, Treatment is defined akin
to Introduction, Methods, etc. from
[DEO, The Discourse Elements Ontology](http://www.sparontologies.net/ontologies/deo/source.html).

```{r http://openbiodiv.net/Treatment}
:Treatment a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement ,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :TaxonomicArticle ] ;
  rdfs:label "Taxonomic Treatment"@en ;
  rdfs:comment "A rhetorical element of a taxonomic publication, where taxon
    			circumscription takes place."@en ;
  rdfs:comment "Таксономично пояснение или само Пояснение е риторчна част
                от таксономичната статия, където се случва описанието
                на дадена таксономична концепция."@bg .
```

##### Class: *Nomenclature*

A taxonomic nomenclature section, or simply a nomenclature, is a rhetorical
element of a taxonomic publication where nomenclatural acts are published and
nomenclatural statements are made. Nomenclature is a subsection of Treatment.

```{r http://openbiodiv.net/NomenclatureSection}
:NomenclatureSection a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :Treatment ] ;
  rdfs:label "Nomenclature Section"@en ;
  rdfs:comment "A section of a taxonomic treatment, containing the scientific
    name of the taxon described by the treatment, and citations to previous
    descriptions, designations of type-genus, and type-species,
    and other information."@en .
```

##### Class: *Nomenclature Heading*

Inside the taxonomic nomenclature section, we have the treatment title.

```{r http://openbiodiv.net/NomenclatureHeading}
:NomenclatureHeading a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement ,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :NomenclatureSection ] ;
                  rdfs:label "Nomenclature Heading"@en ;
  rdfs:comment "Inside the taxonomic nomenclature section, we have the treatment
    title (name of the taxon)."@en .
```

##### Class: *Nomenclature Citations List*

Inside the taxonomic nomenclature section, we have a list of citations.

```{r http://openbiodiv.net/NomenclatureCitationsList}
:NomenclatureCitationsList a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement ,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :NomenclatureSection ] ;
                  rdfs:label "Nomenclature Citations List"@en ;
  rdfs:comment "A section in a treatment that includes the citation of one or
    several previous treatments of the taxon."@en .                  
```

##### Class Definition: *Biology Section*

Subsection of treatment.

```{r http://openbiodiv.net/BiologySection}
:BiologySection a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement, 
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :Treatment ] ;
  rdfs:label "Biology Section"@en .
```

##### Class Definition: *Description Section*

Subsection of treatment.

```{r http://openbiodiv.net/DescriptionSection}
:DescriptionSection a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :Treatment ] ;
  rdfs:label "Description Section"@en .
```

##### Class Definition: *Diagnosis*

Subsection of treatment.

```{r http://openbiodiv.net/DiagnosisSection}

:DiagnosisSection a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement, 
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :Treatment ] ;
  rdfs:label "Diagnosis Section"@en .
```

##### Class: *Distribution Section*

Subsection of treatment.

```{r http://openbiodiv.net/DistributionSection}
:DistributionSection a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement, 
                  [ rdf:type owl:Restriction;
                    owl:onProperty :isContainedBy;
                    owl:someValuesFrom :Treatment ];
  rdfs:label "Distribution Section"@en.
```

##### Class: *Etymology Section*

Subsection of treatment.

```{r http://openbiodiv.net/EtymologySection}
:EtymologySection a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :Treatment ] ;
  rdfs:label "Etymology Section"@en .
```

##### Class: *Taxonomic Key*

Subsection of treatment.

```{r http://openbiodiv.net/TaxonomicKey}
:TaxonomicKey a owl:Class;
  rdfs:subClassOf deo:DiscourseElement, 
                  [ rdf:type owl:Restriction;
                    owl:onProperty :isContainedBy;
                    owl:someValuesFrom :TaxonomicArticle ];
  rdfs:label "Taxonomic Key"@en.
```

##### Class: *Materials Examined*

Subsection of treatment.

```{r http://openbiodiv.net/MaterialsExamined}
:MaterialsExamined a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement ,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :Treatment ] ;
  rdfs:label "Materials Examined"@en .
```

##### Class: *Taxonomic Checklist*

A section in a taxonomic article.

```{r http://openbiodiv.net/TaxonomicChecklist}
:TaxonomicChecklist a owl:Class ;
  rdfs:subClassOf deo:DiscourseElement ,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :isContainedBy ;
                    owl:someValuesFrom :TaxonomicArticle ] ;
  rdfs:label "Taxonomic Checklist"@en ;
  rdfs:comment "A section in a taxonomic article."@en.
```

##### Class: *Taxonomic Name Usage*

In the text of taxonomic articles we find strings like "*Heser stoevi*
Deltschev, sp. n.". In our conceptualization these are called *taxonomic name
usages* (TNU's) as they refer to  published scientific names from the domain
of biological systematics. The taxonomic name usage consists of three parts:

1. Taxonomic name: one or more words identifying the taxon (a purported
evolutionary entity in nature). These can be Latinized or take the form of an
identifier.

2. Taxonomic name author. The name-and-year of the author(s) of the taxonomic name.

3. Taxonomic name status containing information about the type of the taxonomic
name usage.

In the example, "*Heser stoevi*" is the binomial Latinized species name,
"Deltschev" is the name of the person who newly coined the name describing the
taxon , and "sp. n." bears taxonomic (and nomenclatural) information
indicating that this is a species new to science.

Modeling-wise, we consider TNU's to be specialized instances of `pext:Mention`
from the [PROTON Extensions module](http://ontotext.com/proton/). Furthermore,
we link the TNU's to the scientific name they are symbolizing via
`pkm:mentions`.

A taxonomic name usage is the mentioning of a taxon name or a taxon concept label
(see later) in a text, together with possibly an taxonomic status.

```{r http://openbiodiv.net/TaxonomicNameUsage}
:TaxonomicNameUsage rdf:type owl:Class ;
  rdfs:subClassOf :Mention ;
  rdfs:comment "A taxonomic name usage is the mentioning of a taxonomic name
    (scientific name, taxonomic concept label, etc.), optionally with 
    a taxonomic status."@en ;
  rdfs:label "Taxonomic Name Usage"@en .
```

##### Contributed Property: *taxonomic status*
```{r http://rs.tdwg.org/dwc/iri/taxonomicStatus}
dwciri:taxonomicStatus rdf:type owl:ObjectProperty ; 
  rdfs:label "taxonomic status"@en ;
  rdfs:comment "the IRI version of the DwC term taxonomic status"@en .
```

In the logic of our algorithms, it is very important that TNU's are
dated with `dcelements:date`.


### Semantic Model of Biological Nomenclature

In OpenBiodiv, we reify biological names.

In our conceptualization, taxa in nature are things (referents) that are
refered to by our thoughts, theories and concepts (references) that are
labeled or symbolized by by  biological names
([semiotic triangle](https://de.wikipedia.org/wiki/Semiotisches_Dreieck)).

Biological names play a dual role, however, in our system as they are also
concepts, i.e. references of taxonomic name usages. A biological name may
symbolize more than one taxon concept. It is useful to think of biological
names then as taxon concept lineages. More about taxon concepts later.
 
Biological names have been modeled elsewhere such as for example in
[NOMEN](https://github.com/SpeciesFileGroup/nomen). However, NOMEN takes the
approach of using non-human-readable identifiers and only relying on labels to
identify classes of taxonomic names, which does not fit our workflow. For
example, the identifier for the class "biological name" is `NOMEN_0000030`. In
our workflow both RDF generation and debugging would be severely hampered by
this convention. Therefore, OpenBiodiv names use human-readable identifiers.

##### Class : *Taxonomic Name*

```{r http://openbiodiv.net/TaxonomicName}
:TaxonomicName rdf:type owl:Class ;
	skos:prefLabel "Taxonomic Name"@en ;
  skos:altLabel "Biological Name"@en ;
  owl:equivalentClass nomen:NOMEN_0000030 .
```

##### Class: *Scientific Name*

```{r http://openbiodiv.net/ScientificName}
:ScientificName rdf:type owl:Class ;
    rdfs:subClassOf :TaxonomicName ;
    rdfs:label "Scientific Name"@en ;
    owl:equivalentClass  nomen:NOMEN_0000036 .
```

##### Class: *Latinized Scientific Name*

```{r http://openbiodiv.net/LatinName}
:LatinName rdf:type owl:Class ;
    rdfs:subClassOf :ScientificName ;
    skos:prefLabel "Latinized Scientific Name"@en ;
    skos:altLabel "Linnaean Name"@en .
```

##### Class: *Taxonomic Concept Label*

We further introduce the class of taxonomic concept labels, unknown to NOMEN
that is a biological name plus a reference to its description, i.e. it is the
label of taxonomic concept. A taxonomic concept label is a taxonomic name
usage accompanied by an additional part, consisting of "sec." + an identifier
or a literature reference of a work containing the expression of a taxonomic
concept (for example a treatment).*

```{r http://openbiodiv.net/TaxonomicConceptLabel}
:TaxonomicConceptLabel rdf:type owl:Class ;
  rdfs:subClassOf :LatinName ;
  rdfs:label "Taxon Concept Label"@en ;
  rdfs:comment "A taxon concept label is a taxonomic name
    usage accompanied by an additional part, consisting of 'sec.' + an identifier
    or a literature reference of a work containing the expression of a taxon concept
    (treatment)."@en .
```

#### Class: *Operational Taxonomic Unit Identifier*

```{r http://openbiodiv.net/OTU_Id}
:OTU_Id rdf:type owl:Class ;
  rdfs:subClassOf :ScientificName ;
  rdfs:label "Operational Taxonomic Unit Identifier"@en ;
  rdfs:comment "The identifier of an Operational Taxonomic Unit, such as
    BOLD BIN or a Unite SH ID"@en .
```

#### Class: *Vernacular Name*

```{r http://openbiodiv.net/VernacularName}
:VernacularName a owl:Class ;
  rdfs:subClassOf :TaxonomicName ;
  rdfs:label "Vernacular Name"@en ;
  owl:equivalentClass nomen:NOMEN_0000037 .
```

#### Properties

We do not model scientific names down to the level of the Codes as NOMEN does.
For example we do not make a distinction between a zoological and a botanical
name. Nothing prevents us, however, from creating derived classes later on.
This means that our model is somewhat less granular but compatible with NOMEN.

For properties of biological names we take a different path from NOMEN. We
also use different sets of properties to define relationships between
biological names and for their data properties.

For data properties we use DwC terms.

To connect different biological objects such as taxon concepts or occurrences
to a scientific name we use `:scientificName`, which is derived from
`dwciri:scientificName`. Even though `dwciri:scientificName` is defined  in
spirit in
<http://rs.tdwg.org/dwc/terms/guides/rdf/index.htm#2.5_Terms_in_the_dwciri:_namespace>,

To the best of our knowledge, no formal definition of `dwciri:scientificName` exists in RDF.
Therefore, it has been introduced it here together with a super-property to refer to a more broader
class of names.

##### Contributed Property: *name according to*

Contributed property from DwC-IRI.

```{r http://rs.tdwg.org/dwc/iri/nameAccordingTo}
dwciri:nameAccordingTo rdf:type owl:ObjectProperty ;
  rdfs:label "name according to"@en ;
  rdfs:comment "the IRI version of dwc:nameAccordingTo"@en .
```

##### Property: *OpenBiodiv name according to*

```{r http://openbiodiv.net/nameAccordingTo}
:nameAccordingTo rdf:type owl:ObjectProperty ;
  rdfs:subPropertyOf dwciri:nameAccordingTo ;
  skos:prefLabel "OpenBiodiv name according to"@en ; 
  skos:altLabel "sensu"@en ;
  rdfs:range frbr:Expression ;
  rdfs:domain :TaxonomicConceptLabel ;
  rdfs:comment "The reference to the source in which the specific taxonomic 
    concept circumscription is defined or implied - traditionally signified by
    the Latin 'sensu' or 'sec.'' (from secundum, meaning 'according to').
    For taxa that are relevantly circumscribed by identifications, a reference
    to the keys, monographs, experts and other sources should be given. Should
    only be used with IRI's"@en .
```

For relationships between names we introduce two types of relationships:
unidirectional and bidirectional.

##### Property Definition: *has related name'*

 is an object property that we use in order to indicate that two biological
names are related somehow. This relationship is purposefully vague as to
encompass all situations where two biological names co-occur in a text. It is
transitive and reflexive.*

```{r http://openbiodiv.net/relatedName}
:relatedName rdf:type owl:ObjectProperty, owl:TransitiveProperty, owl:ReflexiveProperty ;
  rdfs:label "has related name"@en ;
  rdfs:domain :TaxonomicName ;
  rdfs:range :TaxonomicName ;
  rdfs:comment "'has related name' is an object property that we
    use in order to indicate that two taxonomic names are related somehow. This
    relationship is purposely vague as to encompass all situations where two
    taxonomic names co-occur in a text. It is transitive and reflexive."@en.
```

##### Property Definition: *has replacement name*

This is a uni-directional property. Its meaning is that one one biological
name links to a different biological name via the usage of this property, then
the object of the triple is the form of the biological name the use of which
is more accurate and should be preferred given the information that system
currently holds. This property is only defined for scientific names.*

```{r http://openbiodiv.net/replacementName}
:replacementName rdf:type owl:ObjectProperty ,
                          owl:TransitiveProperty ;
  rdfs:label "has replacement name"@en ;
  rdfs:domain :LatinName ;
  rdfs:range :LatinName ;
  rdfs:comment "This is a uni-directional property. Its meaning
    is that one Linnaean name links to a different Linnaean name via the
    usage of this property, then the object name is more accurate and should be
    preferred given the information that system currently holds. This property is only
    defined for Linnaean names."@en.
```

### Taxonomic Concepts

#### Class Definition: *Taxonomic Concept*

Our view of taxon concepts is based on [Berendsohn
(1995)](http://www.jstor.org/stable/1222443) and [Franz et al
(2008)](http://dx.doi.org/10.1201/9781420008562.ch5).

We consider any given taxon concept to be a scientific theory (concept) about
a class of biological organisms (taxon). The class description, as in "as in
all spiders have spinnerets (silk-producing glands)" (Nico Franz, personal
correspondence), is called *intensional meaning,* whereas the group of
organisms in nature conforming with the intensional meaning is called the
class *extension.*

We want to model both the intensional meaning (which traits do organisms
belonging to a taxon have) and the extension of taxon concepts (which
organisms belong to a taxon) and the extensions being organisms that are
considered to be members of the class.

This necessitates the view that taxon concepts are both instances of a taxon
concept class and are classes of ogranisms. Later, we will show that this
means that we model Taxon Concepts with OWL Full.

OpenBiodiv taxon concepts are instances of `dwc:Taxon` and vice versa (*"A
group of organisms [sic] considered by taxonomists to form a homogeneous
unit."*).

Also, taxon concepts are instances of `frbr:Work` as well, but not vice versa
(*"A distinct intellectual or artistic creation. A work is an abstract entity;
there is no single material object one can point to as the work. We recognize
the work through individual realizations or expressions of the work, but the
work itself exists only in the commonality of content between and among the
various expressions of the work. When we speak of Homer's Iliad as a work, our
point of reference is not a particular recitation or text of the work, but the
intellectual creation that lies behind all the various expressions of the
work."*).

Furthermore, taxon concepts can also be modeled as `skos:Concept`, but not
vice versa (*"A SKOS concept can be viewed as an idea or notion; a unit of
thought. However, what constitutes a unit of thought is subjective, and this
definition is meant to be suggestive, rather than restrictive."*).

All three classes represent a distinctive view that we want to adopt in
modeling different features of taxon concepts.

Holding the views of Berendsohn and of Franz, we require that each taxon
concept is linked to both a biological name and to a work (i.e. publication,
database, etc.), where the circumscription is properly defined.

##### Contributed Class: *Taxon*

```{r http://rs.tdwg.org/dwc/terms/Taxon}
<http://rs.tdwg.org/dwc/terms/Taxon> rdf:type owl:Class ;
                                     dcelements:description "Example: The genus Truncorotaloides as published by Brönnimann et al. in 1953 in the Journal of Paleontology Vol. 27(6) p. 817-820."@en ;
                                     dcelements:hasVersion <http://rs.tdwg.org/dwc/terms/history/#Taxon-2014-10-23> ;
                                     dcelements:issued "2008-11-19" ;
                                     dcelements:modified "2014-10-23" ;
                                     dcelements:replaces <http://rs.tdwg.org/dwc/terms/Taxon-2009-09-21> ;
                                     dwcattributesns:abcdEquivalence "no simple equivalent in ABCD" ;
                                     dwcattributesns:decision <http://rs.tdwg.org/dwc/terms/history/decisions/#Decision-2014-10-26_15> ;
                                     dwcattributesns:status "recommended" ;
                                     rdfs:comment "A group of organisms (sensu http://purl.obolibrary.org/obo/OBI_0100026) considered by taxonomists to form a homogeneous unit."@en ;
                                     rdfs:isDefinedBy <http://rs.tdwg.org/dwc/terms/> ;
                                     rdfs:label "Taxon"@en .
```

##### Class: *Operational Taxonomic Unit*

```{r http://openbiodiv.net/OperationalTaxonomicUnit}
:OperationalTaxonomicUnit rdf:type owl:Class ;
  rdfs:subClassOf skos:Concept, fabio:Work ;
  rdfs:label "Operational Taxonomic Unit"@en ;
  rdfs:comment "A superclass for all kinds of taxonomic hypothesis"@en .
```

##### Class: *Taxonomic Concept*

```{r http://openbiodiv.net/TaxonomicConcept}
:TaxonomicConcept rdf:type owl:Class ;
  owl:equivalentClass dwc:Taxon ;
  rdfs:subClassOf :OperationalTaxonomicUnit ,
                  [ rdf:type owl:Restriction ;
                    owl:onProperty :taxonomicConceptLabel ;
                    owl:minCardinality "1" ] ;
  rdfs:comment "A taxonomic concept in the sense of Berendsohn"@en .
```

##### Property: *has scientific name*

Contributed property from DwC-IRI.

```{r http://rs.tdwg.org/dwc/iri/scientificName}
dwciri:scientificName rdf:type owl:ObjectProperty ;
  rdfs:label "has scientific name"@en;
  rdfs:comment "the IRI version of dwc:scientificName"@en .
```

##### Property: *has OpenBiodiv taxonomic name*

```{r http://openbiodiv.net/taxonomicName}
:taxonomicName rdf:type owl:ObjectProperty ;
  rdfs:label "has OpenBiodiv taxonomic name"@en ;
  rdfs:domain :TaxonomicConcept ;
  rdfs:range :TaxonomicName .
```

##### Property: *has OpeonBiodiv scientific name*

```{r http://openbiodiv.net/scientificName}
:scientificName rdf:type owl:ObjectProperty ;
  rdfs:subPropertyOf :taxonomicName, dwciri:scientificName ;
  rdfs:label "has OpeonBiodiv scientific name"@en ; 
  rdfs:domain :TaxonomicConcept ;
  rdfs:range :ScientificName .
```

#### Property Definition: *has vernacular name*

```{t http://openbiodiv.net/vernacularName}
:vernacularName rdf:type owl:ObjectProperty ;
  rdfs:subPropertyOf :taxonomicName ;
  rdfs:label "has OpenBiodiv vernacular name" @en ;
  rdfs:domain :TaxonomicConcept ;
  rdfs:range :VernacularName .
```

##### Property Definition: *has taxonomic concept label*

```{r http://openbiodiv.net/taxonomicConceptLabel}
:taxonomicConceptLabel rdf:type owl:ObjectProperty ;
  rdfs:subPropertyOf :scientificName ;
  rdfs:label "has taxon concept label"@en ;
  rdfs:domain :TaxonomicConcept ;
  rdfs:range :TaxonomicConceptLabel .
```

##### Property: *has operational taxonomic unit identifier*

```{r http://openbiodiv.net/has_OTU_id}
:has_OTU_Id rdf:type owl:ObjectProperty ;
  rdfs:subPropertyOf :scientificName ;
  rdfs:label "has OTU Id"@en ;
  rdfs:domain :TaxonomicConcept ;
  rdfs:range :OTU_Id .
```

#### Relationships between concepts

##### Class: *RCC5 Relation Dictionary*

A class for all kinds of RCC5 term vocabularies.

```{r http://openbiodiv.net/RCC5RelationshipTerms}
:RCC5Dictionary rdf:type owl:Class ;
  rdfs:subClassOf skos:ConceptScheme ,
                                [ rdf:type owl:Restriction ;
                                  owl:onProperty fabio:isSchemeOf ;
                                  owl:allValuesFrom :RCC5Relation] ;
  rdfs:label "Dictionary of RCC5 Relations"@en .
```

##### Class: *RCC5 Relation*

```{r http://openbiodiv.net/RCC5Relation}
:RCC5Relation rdf:type owl:Class ;
  rdfs:subClassOf skos:Concept ;
  rdfs:subClassOf [ rdf:type owl:Restriction ;
                    owl:onProperty <http://www.w3.org/2004/02/skos/core#inScheme> ;
                    owl:someValuesFrom :RCC5Dictionary ] ;
  rdfs:label "RCC5 Relation"@en ;
  rdfs:comment "The of RCC 5 relation, e.g. 'partially overlaps'"@en .
```

##### Class: *RCC5 Statement*

```{r http://openbiodiv.net/RCC5Statement}
:RCC5Statement rdf:type owl:Class ;
  rdfs:subClassOf ptop:Statement ;
  rdfs:label "RCC5 Statement" ;
  rdfs:comment "A statement of RCC-5 relationship" .
```

##### Property: *generic RCC5 Property*

```{r http://openbiodiv.net/rcc5Property}
:rcc5Property rdf:Type owl:ObjectProperty ;
  rdfs:domain :RCC5Statement .
```

##### Property: *RCC5 from region*

```{r http://openbiodiv.net/rcc5fromRegion}
:rcc5fromRegion rdf:type owl:ObjectProperty ;
  rdfs:subPropertyOf :rcc5Property ;
  rdfs:label "from region" ;
  rdfs:comment "Connects to the RCC5 statement to the originating region"@en .
```

##### Property: *RCC5 to region*

```{r http://openbiodiv.net/rcc5toRegion}
:rcc5toRegion rdf:type owl:ObjectProperty ;
  rdfs:label "to region" ;
  rdfs:subPropertyOf :rcc5Property ;
  rdfs:comment "Connects to the RCC5 statement to the target region"@en .
```

##### Property: *RCC5 relation type* 

```{r http://openbiodiv.net/rcc5RelationType}
:rcc5RelationType rdf:type owl:ObjectProperty ;
  rdfs:label "relation type" ;
  rdfs:subPropertyOf :rcc5Property ;
  rdfs:range :RCC5Relation ;
  rdfs:comment "Connects an RCC-5 statement to the type of RCC-5 relation between the regions."@en .
```

##### Property: *RCC5 statement according to* 

```{r http://openbiodiv.net/rcc5AccordingTo}
:rcc5AccrodingTo rdf:type owl:ObjectProperty ;
  rdfs:subPropertyOf :rcc5Property ;
  rdfs:label "RCC5 statement according to" ;
  rdfs:range frbr:Expression ;
  rdfs:comment "In which publication was the statement made"@en .
```

In order to model complex RCC-5 statements such as those made in
[Jansen & Franz 2015](https://doi.org/10.3897/zookeys.528.6001) we use 
the above defitions and the RCC5 Vocabulary described in the Appendix.


## Vocabularies

### Vocabulary of Taxonomic Statuses

[Taxonomic name usages](#taxonomic-name-usage) (TNU's) in taxonomic articles
may be accompanied by strings such as "new. comb.", "new syn.", "new record
for Cuba", and many others. These postfixes to a taxonomic name usage are
called in our model taxonomic statuses and have taxonomic as well
nomenclatural meaning. For example, if we are describing a species new to
science, we may write "n. sp." after the species name, e.g. "*Heser stoevi*
Deltchev sp. n." This particular example is also a nomenclatural act in the
sense of the Codes of zoological or botanical nomenclature.

Not all statuses are necessarily nomenclatural in nature. Sometimes the status
is more of a note to the reader and conveys taxonomic rather than
nomenclatural information. E.g. when a previously known species is recorded in
a new location.

Here we take the road of modeling statuses from the bottom-up, i.e. based on
their actual use in three of the most successful journals in biological
systematics - ZooKeys, Biodiversity Data Journal, and PhytoKeys. We have
analyzed about 4,000 articles from these journals and have come up with a
vocabulary of statuses described below. The terms in this vocabulary refer to
broad concepts and encompass both specific cases of botanical or zoological
nomenclature as well as purely taxonomic or informative use. We believe these
concepts to be adequately granular for the purposes of reasoning in OpenBiodiv
(see the paragraphs on rules in [biological names](#biological-names).

## Extraction of 4,000 TNU suffixes

1. The first step we took in creating this vocabulary is to extract the actual
real-world taxonomic statuses as found almost 4,000 Pensoft articles. The
script to achieve this is found under [mine_tnu.R](R/mine_tnu.R). The output of
the script is a list of taxonomic status abbreviations as a text file,
[taxonomic_statuses.txt](status_abbrev/all.txt), in total 253 distinct
usages.

2. The second step we took is to normalize these statuses manually: clean up
parts of abbreviations that are not part of the status (such as particular
taxon names) or remove irrelevant statuses (such as geological information).

3. After cleaning up the usages we came up with the terms in the next section
and assigned each of remaining status abbreviations to one of the terms. The
assignments can be seen in text files with correspondning names.

For a similar attempt, see
<http://rs.gbif.org/vocabulary/gbif/taxonomic_status.xml>. We do map our
statuses to the GBIF statuses to the best of our knowledge but we have more
statuses.

##### Class: *Taxonomic Status Dictionary*

```{r http://openbiodiv.net/TaxonomicStatusDictionary}
:TaxonomicStatusDictionary rdf:type owl:Class ;
  rdfs:subClassOf <http://www.w3.org/2004/02/skos/core#ConceptScheme> ,
                                [ rdf:type owl:Restriction ;
                                  owl:onProperty fabio:isSchemeOf ;
                                  owl:allValuesFrom :TaxonomicStatus] ;
  rdfs:label "Dictionary of Taxonomic Statuses" ;
  fabio:hasDiscipline <http://dbpedia.org/page/Taxonomy_(biology)> .
```

##### Class: *Taxonomic Status Term*

```{r http://openbiodiv.net/TaxonomicStatus}
:TaxonomicStatus rdf:type owl:Class ;
  rdfs:subClassOf skos:Concept ;
  rdfs:subClassOf [ rdf:type owl:Restriction ;
                    owl:onProperty <http://www.w3.org/2004/02/skos/core#inScheme> ;
                    owl:someValuesFrom :TaxonomicStatusDictionary ] ;
  rdfs:label "Taxonomic Status Term"@en ;
  rdfs:comment "The status following a taxonomic name usage in a taxonomic
                manuscript, i.e. 'n. sp.',
                'comb. new',
                'sec. Franz (2017)', etc"@en .
```

