##### Vocabulary: *OpenBiodiv Taxonomic Status Vocabulary*

```{r http://openbiodiv.net/TaxonomicStatusTerms}
:TaxonomicStatusTerms a :TaxonomicStatusDictionary ;
 rdfs:label "OpenBiodiv Vocabulary of Taxonomic Statuses"@en ;
```

##### Term: *Taxonomic Uncertainty*

When a TNU is followed by the term `:TaxonomicUncertainty`, the implication is
that the taxonomic concept identified by the name has an uncertain placement in
the biological taxonomy, or if we are talking about a specimen or a sample of
some kind that we are unable to identify the taxonomic down to its rank. This
is related to <http://rs.gbif.org/vocabulary/gbif/taxonomicStatus/accepted>

```{r http://openbiodiv.net/TaxonomicUncertainty}
:TaxonomicUncertainty a :TaxonomicStatus ;
  rdfs:label "Taxonomic Uncertainty"@en ;
  skos:inScheme :TaxonomicStatusTerms ;
  skos:related <http://rs.gbif.org/vocabulary/gbif/taxonomicStatus/accepted> ;
 
  rdfs:comment "The implication of this term that the taxonomic concept identified
by a name has an uncertain placement in the biological taxonomy, or if we
are talking about a specimen or a sample of some kind that we are unable to
identify the taxonomic down to its rank. "@en .
```

Here're some ways in which it can be abbreivated:

[taxonomic_uncertainty.txt](status_abbrev/taxonomic_uncertainty.txt)

##### Term: *Taxonomic Discovery*

When a TNU is followed by the term `:TaxonDiscovery`, the implication is that
the present context in which the TNU is used is circumscribing a new taxon
concept (a taxon concept of a taxon new to science), and simultaneously
assigning it a new name.

```{r http://TaxonomicDiscovery}
:TaxonomicDiscovery a :TaxonomicStatus ;
  skos:inScheme :TaxonomicStatusTerms ;
  rdfs:label "Taxon Discovery"@en ;

  rdfs:comment "When a TNU is followed by the term `:TaxonDiscovery`, the
implication is that the present context in which the TNU is used is
circumscribing a new taxon concept (a taxon concept of a taxon new to
science), and simultaneously assigning it a new name."@en .
```

Here're some ways in which it can be abbreivated:

[taxon_discovery.txt](R/taxon_discovery.txt)

##### Term: *Replacement Name*

When a TNU is followed by the term `:ReplacementName`, the implication is that
the name that is referred to by the TNU is replacing another name for various
reasons, and thus becoming the preferred/accepted/available way of referring to
whatever taxonomic concepts the now replaced name had been referring to. Cases
include changes of rank, new combinations, spelling mistakes, etc.

```{r http://openbiodiv.net/ReplacementName}
:ReplacementName a :TaxonomicStatus ;
  skos:inScheme :TaxonomicStatusTerms ;
  rdfs:label "Replacement Name"@en ;
  rdfs:comment "When a TNU is followed by the term `:ReplacementName`, the implication is that
the name that is referred to by the TNU is replacing another name for various
reasons, and thus becoming the preferred/accepted/available way of refering to
whatever taxon concepts the now replaced name had been referring to. Cases
include changes of rank, new combinations, spelling mistakes, etc.
"@en .
```
Here're some ways in which it can be abbreviated:

[replacement_name.txt](R/replacement_name.txt)

##### Term: *Unavailable Name*

When a TNU is followed by the term `:UnavailableName`, the implication is that
the name that is being referred to by the TNU is no longer or has never been
available for use due to the fact that it has either been replaced or it has
been determined that the name has been improperly coined or published, or the
name contains any general error rendering it unfit for use. This is the
same as <http://rs.gbif.org/vocabulary/gbif/taxonomicStatus/synonym>

```{r http://openbiodiv.net/UnavailableName}
:UnavailabeName a :TaxonomicStatus ;
  skos:inScheme :TaxonomicStatusTerms ;
  rdfs:label "UnavailableName"@en ;
  skos:exactMatch  <http://rs.gbif.org/vocabulary/gbif/taxonomicStatus/synonym> ;
  rdfs:comment "When a TNU is followed by the term `:UnavailableName`, the implication is that
the name that is being referred to by the TNU is no longer or has never been
available for use due to the fact that it has either been replaced or it has
been determined that the name has been improperly coined or published, or the
name contains any general error rendering it unfit for use."@en .
```

Here are some ways in which it can be abbreviated:

[unavaible_name.txt](R/unavavaible_name.txt)

##### Term: *Available Name*

When a TNU is followed by the term `:AvailableName`, the implication is that
the name that is being referred to by the TNU has been determined to be fit
for use either by revoking an older act rendering the name unavailable or by
fixing other issues with the name or finding out that other issues with the
name had been fixed, or just stating the fact that the name shall be used or
even conserving it, so that the name can be freely used from then on in
compliance with all Codes and practices. This is the same as
<http://rs.gbif.org/vocabulary/gbif/taxonomicStatus/accepted>

```{r http://openbiodiv.net/AvaialbleName}
:AvailableName a :TaxonomicStatus ;
  skos:inScheme :TaxonomicStatusTerms ;
  rdfs:Label "Available Name"@en ;
  skos:exactMatch <http://rs.gbif.org/vocabulary/gbif/taxonomicStatus/accepted> ;
  rdfs:comment "When a TNU is followed by the term `:AvailableName`, the
implication is that the name that is being referred to by the TNU has been
determined to be fit for use either by revoking an older act rendering the
name unavailable or by fixing other issues with the name or finding out that
other issues with the name had been fixed, or just stating the fact that the
name shall be used or even conserving it, so that the name can be freely used
from then on in compliance with all Codes and practices."@en.
```

[available_name.txt](R/available_name.txt)

##### Term: *Type Species Designation*

When a TNU is followed by the term `:TypeSpeciesDesignation`, the implication
is that the taxon concept of the name in the TNU as understood by the author
should be considered the type species of a higher taxon as understood by the
author.

```{r http://openbiodiv.net/TypeSpeciesDesignation}
:TypeSpeciesDesignation a :TaxonomicStatus ;
  rdfs:label "Type Species Designation" @en ;
  skos:inScheme :TaxonomicStatusTerms ;
  rdfs:comment "When a TNU is followed by the term `:TypeSpeciesDesignation`, the implication
is that the taxon concept of the name in the TNU as understood by the author
should be considered the type species of a higher taxon as understood by the
author."@en.
```

[type_species_designation.txt](R/type_species_designation.txt)

##### Term: *Type Specimen Designation*

When a TNU is followed by the term `:TypeSpecimenDesignation`, the implication
is that the specimen identified by the name in TNU should be considered a type
of the taxonomic concept identified by the name as understood by the author of
TNU.

```{r http://openbiodiv.net/TypeSpecimenDesignation}
:TypeSpecimenDesignation a :TaxonomicStatus ;
  rdfs:label "Type Specimen Designation" @en ;
  skos:inScheme :TaxonomicStatusTerms ;
  rdfs:comment "When a TNU is followed by the term `:TypeSpecimenDesignation`, the implication
is that the specimen identified by the name in TNU should be considered a type
of the taxon concept identified by the name as understood by the author of
TNU."@en.
```

[type_specimen_designation.txt](R/type_specimen_designation.txt)

##### Term: *New Occurrence Record*

When a TNU is followed by the term `:NewOccurrenceRecord`, the implication is that the
description of taxon concept of the name as understood by the author is being
extended with new occurrences (for a given locality).

```{r http://openbiodiv.net/NewOccurrenceREcord}
:NewOccurrenceRecord a :TaxonomicStatus ;
  rdfs:label "New Occurrence Record" @en ;
  skos:inScheme :TaxonomicStatusTerms ;
  rdfs:comment "When a TNU is followed by the term `:NewRecord`, the implication is that the
description of taxon concept of the name as understood by the author is being
extended with new occurrences (for a given locality).
"@en.
```

[new_occurrence_record.txt](R/new_occurrence_record.txt)
