---
title: "R Demo for TDWG Conference 2016"
author: "Viktor Senderov"
date: "`r Sys.Date()`"
---


## Summary

This is the R demo for the TDWG conference. 

## Demo

`rdf4jr` library for communication with the GraphDB server.
`obkms` the OBKMS library for RDF-ization and query.

```{r , purl=TRUE, eval = FALSE}
library(obkms)
library(rdf4jr)
```

Read-in a sample XML file.

```{r , purl=TRUE, eval = FALSE}
example = "/media/viktor/www2/main/vsenderov.backup/SRC/TaxPub2RDF/BDJ_Archive/10.3897_BDJ.4.e8150.xml"
configuration_file = "/media/viktor/www2/main/vsenderov.backup/SRC/TaxPub2RDF/etc/config.yml"
taxpubs = command_line_args[ 2:length( command_line_args ) ]
server_access_options = yaml::yaml.load_file( configuration_file )
taxpubs = c("/media/viktor/www2/main/vsenderov.backup/SRC/TaxPub2RDF/Standard_Model_Conversions/10.3897-BDJ.4.e10095.xml", 
            "/media/viktor/www2/main/vsenderov.backup/SRC/TaxPub2RDF/Standard_Model_Conversions/bdj4701.xml")
```

Download BDJ articles
http://bdj.pensoft.net/lib/journal_archive.php?journal=bdj&date=26%2F3%2F2015
[3:01:12 PM] Teodor Georgiev: returned all articles after date
http://bdj.pensoft.net/lib/journal_archive.php?journal=bdj&date=26%2F3%2F2015
where {date} is a urlencoded date in format day/month/year 
(e.g. 26/3/2014)
[3:01:26 PM] vsenderov: got it
[3:01:27 PM] Teodor Georgiev: 
If date is not specified - all articles from the specified journal are 
returned.
TODO: zoobank thing
look at 10.3897_BDJ.4.e10095.xml:          <tp:taxon-name><tp:taxon-name-part taxon-name-part-type="genus">Heser</tp:taxon-name-part> <tp:taxon-name-part taxon-name-part-type="species">stoevi</tp:taxon-name-part><object-id content-type="zoobank" xlink:type="simple" object-id-type="ZooBank">urn:lsid:zoobank.org:act:E4D7D5A0-D649-4F5E-9360-D0488D73EEE8</object-id></tp:taxon-name>
grep: turtle: Is a directory
TODO taxon authority (e972.xml) nomenclature
TODO  synonyms e6577
TODO Caption extractor from figures and tables
TODO take care of subspieces
```{r}
bdj.archive_folder = "/media/viktor/www2/main/vsenderov.backup/SRC/TaxPub2RDF/BDJ_Archive"
bdj.taxpubs = list.files(bdj.archive_folder)
bdj.processing_status = list()

#taxpubs = taxpubs[round(runif(n= 10, min = 1, max = length(taxpubs)))]
for ( taxpub_file in bdj.taxpubs ) {
  ttl = xml2rdf(paste0( bdj_archive, "/", taxpub_file ) ) 
  processing_status = httr::content( rdf4jr::add_data(server_access_options, server_access_options$repository, ttl), as="text" ) 
  cat ( processing_status )
  bdj.processing_status[[taxpub_file]] = paste0("processed: ", processing_status)

  ttl_name = gsub( ".xml", ".ttl", taxpub_file)
  connex = file( paste0(bdj_archive, "/turtle/", ttl_name ) )
  writeLines(ttl, con = connex )
  close( connex )
}
```

TODO : error "10.3897_BDJ.3.e4162.xml"
TODO: wrong rank of order http://id.pensoft.net/de9902eb-f538-419e-8b27-2a0041154e25

Extract ZooKeys:

```{r}
zookeys.archive = "/media/viktor/www2/main/vsenderov.backup/SRC/TaxPub2RDF/ZooKeys_Archive"
zookeys.taxpubs = list.files(zookeys_archive)
zookeys.status = list()
#taxpubs = taxpubs[round(runif(n= 10, min = 1, max = length(taxpubs)))]
for ( taxpub_file in zookeys.taxpubs ) {
  print( taxpub_file )
  ttl = xml2rdf(paste0( zookeys_archive, "/", taxpub_file ) ) 
  status =  httr::content( rdf4jr::add_data(server_access_options, server_access_options$repository, ttl), as="text")
  cat ( status )
  zookeys.status[[taxpub_file]] = paste0("processed: ", status)
  ttl_name = gsub( ".xml", ".ttl", taxpub_file)
  connex = file( paste0(bdj_archive, "/turtle/", ttl_name ) )
  writeLines(ttl, con = connex )
  close( connex )
}
```

Get the username-password or API key from the environment:

```{r , purl=TRUE, eval = FALSE}
Sys.setenv(OBKMS_SECRET="obkms:1obkms")
server_access_options$userpwd = Sys.getenv("OBKMS_SECRET")
```

Test GraphDB connectivity:
```{r, purl = TRUE, eval = FALSE}
stopifnot( get_protocol_version( options ) == "6" )
```

Read in the XML file:
```{r , purl=TRUE, eval = FALSE}
### currently only one file accepted
#xml = read_xml(taxpubs[2])
xml = XML::xmlParse ( taxpubs[2])
XML::free ( xml)
```

Now, we will write different functions to extract different features of the XML.

```

```

We write a function that returns RDF about the publisher. RDF will be returned as strings in Turtle Syntax.
```
get_publisher_rdf = function(taxpub_xml) {
  publisher_name = xml_text(xml_find_all(taxpub_xml, "/article/front/journal-meta/publisher/publisher-name")[[1]])
  publisher_id = get_or_create_node_by_label( obkms_opts, publisher_name) 
  if (publisher_id == null) publisher_id = mint_new_pensoft_id()
  
  journal_name = xml_text(xml_find_all(taxpub_xml, "/article/front/journal-meta/journal-id")[[1]])
  journal_id = get_or_create_node_by_label( obkms_opts, journal_name)
  
  
  rdf = "
  prefix fabio: <fdfsdsfsdfs> .
  %journal_id a fabio:Journal"
  return(rdf)
}
```

# this is in semi pseudo code
extract_journal_metadata = function(y_document) {
  journal_name = get_xml_value("/article/front/journal-meta/journal-id")
}
