---
title: "Import the CrossRef journal database into GraphDB"
author: "Viktor Senderov"
date: "June 24, 2016"
output: html_document
---

library("rdf4jr")
rdf4j_server_options = create_rdf4j_server_options(server_url = "http://213.191.204.69:7777/graphdb", authentication = "basic_http", userpwd = "obkms:1obkms")
testthat::expect_output(get_protocol_version(rdf4j_server_options), "[0-9]")


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCurl)
library(uuid)

graphDbOptions = list(
  protocol = "https://",
  server = "rdf.s4.ontotext.com/4937448214/OBKMS" ,
  userpwd = "vsenderov:1obkms" ,
  api_key = "s4i0k6o7v6l0" ,
  secret = "9k03ah49e3da607"
)

# acceptable values 
# protocol = "http://", or "https://"
# authentication = "http" or "api"
create_rdf4j_server_options = function (protocol = "http://", server, authentication = "http", userpwd, api_key, secret) {
  if (authentication == "http") {
  return ( list(
    protocol = protocol,
    server = server,
    authentication = authentication,
    userpwd = userpwd
  ) )  
  }
  else {
    return ( list (
      protocol = protocol,
      server = server,
      authentication = "api",
      api_key = api_key ,
      secret = secret
    ))
  }
}
```

## CrossRef Journal Dabase

CrossRef provides a journal database at <http://www.crossref.org/titleList/> .

This script downloads the database, process it, and imports into GraphDB:

```{r echo=TRUE}
crjdataUrl = "http://ftp.crossref.org/titlelist/titleFile.csv"
```

## Download and process the database into a dataframe

```{r echo=TRUE}
crjData.csv = getURL( url = crjdataUrl )
save(file = "crjData.csv", crjData.csv)
remove(crjData.csv)
journals = read.csv( textConnection( crjData.csv ) )
```

## Functions

A set of helper functions to access GraphDB. They all require the GraphDB configurtion in the beginning to be set.

```{r echo = TRUE}
getRdfProtocol = function( options = graphDbOptions , auth = "basic") {
  endpoint = "/protocol"
  if ( auth == "basic" ) {
    getURL( paste(options$protocol, options$server, endpoint, sep = ""), verbose = TRUE, userpwd = graphDbOptions$userpwd, httpauth = 1L)
  }
  else { #API-KEY Authentication
     getURL( paste(options$protocol, options$api_key, ":", options$secret, "@", options$server, endpoint, sep = ""), verbose = TRUE, userpwd = graphDbOptions$userpwd, httpauth = 1L)
  }
}

getRdfRepos = function( options = graphDbOptions ) {
  endpoint = "/repositories"
  r = getURL( paste(options$server, endpoint, sep = ""), verbose = TRUE, userpwd = graphDbOptions$userpwd, httpauth = 1L, httpheader = c(Accept = "application/sparql-results+xml, */*;q=0.5"))
  cols = c("uri", "id", "title", "readable", "writable")
  d = xmlTreeParse(r, asText = TRUE, useInternalNodes = TRUE)
  
  ns = getNodeSet(d, "//x:results/x:result/x:binding[@name = 'uri']/x:uri", "x")
  uri = sapply (ns, xmlValue)
  
  ns = getNodeSet(d, "//x:results/x:result/x:binding[@name = 'id']/x:literal", "x")
  id = sapply (ns, xmlValue)
  
  ns = getNodeSet(d, "//x:results/x:result/x:binding[@name = 'title']/x:literal", "x")
  title = sapply (ns, xmlValue)
  
  ns = getNodeSet(d, "//x:results/x:result/x:binding[@name = 'title']/x:literal", "x")
  title = sapply (ns, xmlValue)
  
  ns = getNodeSet(d, "//x:results/x:result/x:binding[@name = 'readable']/x:literal", "x")
  readable = sapply (ns, xmlValue)
  
  ns = getNodeSet(d, "//x:results/x:result/x:binding[@name = 'writable']/x:literal", "x")
  writable = sapply (ns, xmlValue)
  
  data.frame(uri, id, title, readable, writable)
}

getRdfQuery = function ( options = graphDbOptions, repoId, query, auth = "basic", accept = "application/sparql-results+xml, */*;q=0.5", queryLn = "sparql", infer = "true", varbindings = "", timeout = "0" ) {
  endpoint = paste( options$server, "/repositories/", repoId, sep = "") 
  r = getURL( paste( endpoint, "?query=", URLencode(query), "&queryLn=",queryLn, sep = "") , verbose = TRUE, userpwd = graphDbOptions$userpwd, httpauth = 1L, httpheader = c(Accept = accept))
}

# POSTs a query to rdf4j endpoint
# accept accepts the following values:
# XML
# queryLn
# SPARQL
# SERQL
postRdfQuery = function ( options = RDF4J_OPTIONS, repoId, query, accept = "XML", queryLn = "SPARQL", infer = "true", varbindings, timeout = "0" ) {
  # need to construct a POST query
  # curl -u obkms:1obkms -X POST --header "Accept:application/sparql-results+xml" --data   "query=SELECT%20(COUNT(*)%20as%20?count)%0AFROM%20%3Chttp://www.ontotext.com/implicit%3E%0AWHERE%20%7B%0A%20%20%20?s%20?p%20?o%20.%0A%7D" http://213.191.204.69:7777/graphdb/repositories/OBKMS
  
  # many issues here, see http://stackoverflow.com/questions/5797688/post-request-using-rcurl ,  more specifically comments by Duncan

  url  = paste(options$protocol, options$server, "/repositories/", repoId, sep = "")
  httpheader = c(Accept = "application/sparql-results+xml")
  curl = getCurlHandle()
  # Note the header that extends across two lines with the second line
  # prefixed with white space.
  curlSetOpt( .opts = list(httpheader = c(Accept = "text/csv, */*;q=0.5") ,
  verbose = TRUE), userpwd = options$userpwd, httpauth = 1L,
  curl = curl)
  postForm( url,
            query = query, curl = curl,  style = "POST")
  
}
```

```{r echo = TRUE}
graphDbExec = function ( query, returnFormat = "text/csv", update = FALSE ) {
  # config : change this if you change your database
  endpoint = "http://213.191.204.69:7777/graphdb/sparql"
  updateEndpoint = "http://213.191.204.69:7777/graphdb/update"
  userpwd = "obkms:1obkms"
  
  # exec
  if (update) endpoint = updateEndpoint
  r = getURL(url = paste(endpoint, URLencode(query), sep = "?query="), userpwd = userpwd, httpheader = c(Accept=returnFormat), verbose = TRUE , httpauth= 1L)
}
```

The following function creates nodes for each of the subjects found in the database.

```{r echo=TRUE}
createSubjects = function ( subjects ) {
  sapply ( subjects, function ( s ) {
    # `s` contains a single subject string
    # for example, s = "Peace"
    # this is a how a subject needs to be expressed as triples
    # ```
    # @prefix fabio: <http://purl.org/spar/fabio/>
    # @prefix dc: <http://purl.org/dc/elements/1.1/>
    # <http://id.pensoft.net/subjects/Peace a fabio:SubjectDiscipline ;
    #                                   dc:title "Peace" .
    # ```
    
    # first, need to create an identifier for the discipline
    uuid= UUIDgenerate()
    query = 'PREFIX fabio: <http://purl.org/spar/fabio/>
             PREFIX dc: <http://purl.org/dc/elements/1.1/>
             INSERT DATA {
               <http://id.pensoft.net/subjects/someid> a fabio:SubjectDiscipline ;
                                                          dc:title "sometitle" .
             }'
    query = gsub("someid", uuid, query)
    query = gsub("sometitle", as.character(s), query)
    
    # next, need to execute the query
    r = graphDbExec(query)
    r = getURL(url = paste(endpoint, URLencode(query), sep = "?query="), userpwd = "obkms:1obkms", httpheader = c(Accept="text/csv"), verbose = TRUE , httpauth= 1L)
  })
}
```
